Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages

### 2021-04-26 manuscript/bef_stability/manuscript.Rmd
Updates

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
csl: ecology-letters.csl
---

@@

- Number of references: 92
- Number of figures: 2
- Number of tables: 1
- Number of text boxes: 0

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural context. Here we study the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardized monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first reveal that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we show that food-web structure affects community stability, with a positive
effect of mean trophic level, and contrary to expectation, a destabilizing
effect of species richness. Finally, we find that environmental conditions
affect both biomass and its stability via effects on richness and network
structure. Strikingly, effects on stability are mediated by population stability
rather than synchrony, which contrasts with results from single-trophic
communities and might be characteristic of multi-trophic communities.

@@

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a major effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009], because it largely determines the
amount and the distribution of matter and energy fluxes within and among
ecosystems [@loreau_meta-ecosystems:_2003]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

@@

Species richness has been found to increase primary productivity, community biomass
and its temporal stability in a diversity of ecosystems and taxa both
empirically and experimentally [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020]. However, the relationship between
diversity and ecosystem functioning remains mostly focused on single trophic
levels, generally the primary producers, ignoring the effects of food-web
structure. This gap is even more striking for the relationship between diversity
and temporal stability, in spite of long-standing interest for the links between
food-web complexity and stability in theoretical ecology [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and food-web structure should modulate ecosystem properties such as
community biomass and its stability. Complementarity in the use of
resources is important to understand the effects of species richness on
productivity in food-webs [@thebault_food-web_2003; @poisot_trophic_2013;
@schneider_animal_2016], but it also depends on food-webs connectance, i.e. its
level of generalism [@thebault_food-web_2003; @thebault_cascading_2007]. In
addition, the presence of species at higher trophic levels may increase total
primary production, through possibly through the selection of primary producers
which are larger and more complementary [@wang_biodiversity_2018].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in population temporal variations also explains a significant part
of the stabilizing effects of species richness on community biomass in food-webs
[@thebault_trophic_2005]. Indeed, a low synchrony in population fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a population is compensated by the increase of another. However,
higher food-web connectance can counterbalance the positive effects of species
richness on temporal stability of community biomass, because it may increase
competition between consumers and then leads to a positive relationship between
species richness and population temporal variability
[@thebault_trophic_2005]. The addition of predators has
also cascading effects on both population variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and food-web structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The lack of studies that consider jointly effects of diversity and food-web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food-web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that
describing trophic interactions between species is highly labour intensive.
These challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires longer-term perspectives. In this
context, recent developments to predict species trophic interactions
[@beckerman_foraging_2006; @gravel_inferring_2013; @portalier_mechanics_2019],
combined with observational long-term surveys of ecological communities, can
offer the opportunity to investigate the links between diversity, food-web
structure and the stability of total community biomass in natural settings.

The joint study of the effects of diversity and food web structure on total
biomass and its stability is even more important in the context of environmental
changes. Recent studies on natural communities have shown that environment may
alter community biomass [@grace_integrative_2016; @woods_testing_2020] and its
temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
food-web structure, as larger ecosystems host food-webs with higher diversity
and trophic levels [@mchugh_dual_2010] whereas temperature can affect food-web
diversity and structure in contrasted ways depending on the ecosystem studied
[@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural communities,
the relationships among species richness, community biomass and its temporal
stability might also arise from effects of the environment on both diversity and
ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016]. Thus, the effect of environment on ecosystem
functioning might arise from direct effects or indirectly through effects on
species richness or food-web structure.

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @beckerman_foraging_2006;
@gravel_inferring_2013; @brose_predator_2019; @portalier_mechanics_2019].

Our aim is to document the relationships among species richness, food-web
structure, community biomass and its temporal stability in natural communities. To
do so, we used a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years. We inferred 1496 interaction
networks, based on body size and diet. We question whether species richness,
food-web connectance and average trophic level are related to community biomass and its
temporal stability. We further assess how ecosystem size and temperature
mediated this relationship.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all fishes and the resources (Fig.
SXX), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (appendix B, Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

@@

Species richness and mean trophic level had a significant positive direct effect
on community biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-bm)A, Table
\@ref(tab:indir-sem)), while connectance had no direct effect.

The two PCA axes synthesising environmental variables had no significant direct
effects on community biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, related to the stream size, was positively linked to community biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, related to the average
annual temperature was positively related
to community biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). Streams
with higher temperature and BOD contained more species, indirectly increasing
community biomass. The total effect of PCA1, PCA2 and species richness on
community biomass were of the same magnitude.

@@

By its effects on synchrony and population variability ($\overline{CV_{sp}}$), species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)), which outweighted the decreasing effect of species
richness on synchrony. The
negative effect of $\overline{CV_{sp}}$ on temporal stability was twice higher
than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-stab)A).

The total effect of mean trophic level and species richness on temporal
stability of biomass were of the same magnitude (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-stab)A) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Food webs in which community biomass is located on
average at higher trophic levels thus tend to have higher population and
community stability. Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-stab)A, Table \@ref(tab:indir-sem)).

The total effect of PCA2, related to average annual temperature and BOD, on temporal
stability was of higher magnitude than the effects of species richness and mean
trophic level (Table \@ref(tab:indir-sem)).
The PCA2 had a total negative effect on temporal stability, meaning that higher
temperature and BOD results in lower stability
($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table \@ref(tab:indir-sem)). 
PCA2 had an indirect positive effect on temporal stability through a
direct positive effect on mean trophic level ($r_\delta = 
`r env_on_stab$via_tlvl["log_RC2"]`$, Table \@ref(tab:indir-sem)). However, PCA2
had a negative effect on temporal stability through a direct positive effect on 
$\overline{CV_{sp}}$ and species richness (resp.  $r_\delta = 
`r env_on_cv_sp$direct["log_RC2"]`$ and $r_\delta = `r env_on_stab$via_richness["log_RC2"]`$,
Table \@ref(tab:indir-sem)). This was also true for the PCA1, related to stream
size, with a resulting very slight positive effect of PCA1 on temporal stability
($r_\delta = `r env_on_stab$total["log_RC1"]`$, Table \@ref(tab:indir-sem)).
Warmer and larger streams thus tend to have more species and more variability in
population biomass, then have a lower stability of community biomass.

@@

We found that species richness increased community biomass as already
found in plant and tree communities [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018] as well
as in fish communities [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. The positive effect of species richness on community biomass
might be due to resource use complementarity [@carey_determining_2011], which is
driven by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability to include high biomass
species with higher species richness [@loreau_biodiversity_2001]. Our results
are also in agreement with a model predicting a positive relationship between
animal diversity and total consumer biomass in size-structured food-web
[@schneider_animal_2016].

Surprisingly, we found a negative relationship between temporal stability and
species richness. It contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006; @valencia_synchrony_2020], mesocosms
[@pennekamp_biodiversity_2018] with studies on natural communities such as
butterflies  or lotic fishes [@olivier_urbanization_2020; @franssen_annual_2011
], where positive relationships were found between temporal community
stability and species richness. However, some other studies also found a
negative relationship, for instance in natural grasslands [@yang_effects_2011]
or a marginal relationship in natural forests [@declerck_species_2006].

As most studies on experimental grasslands [@isbell_biodiversity_2009;
@roscher_identifying_2011] and natural communities of birds, grasslands, bats
and butterflies [@bluthgen_land_2016; @olivier_urbanization_2020], we found that
species richness decreases synchrony, thereby increasing stability [but see
@declerck_species_2006; @valencia_synchrony_2020]. Our study thus extends the
evidence that species richness decreases synchrony to the case of multi-trophic
communities. Species richness can reduce synchrony when (i) a given increase in species
richness raises demographic stochasticity by decreasing population
sizes, or when (ii) species have different responses to environmental variation
[@loreau_species_2008].

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
studies focusing on single-trophic communities have reported mixed effects of
species richness on $\overline{CV_{sp}}$, with destabilising effects of species
richness at population level in grasslands [@jiang_different_2009;
@gross_species_2014], a previous meta-analysis suggested that species richness had
often a positive effect on population stability in multi-trophic communities
[@jiang_different_2009].
In food-webs, theoretical models predict that the relationship between temporal
stability of populations and species richness depends on interaction strength
and its distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@downing_temporal_2020], with stronger interactions promoting a negative
relationship between species richness and stability. Therefore, it is possible
that our results are due to strong trophic interactions, leading to more
fluctuations in population biomass, but this hypothese was not testable with the
available dataset.  Alternatively, our results might be explain by the effect of
increasing species richness which can increase demographic stochasticity.

@@

The average trophic level was positively linked to community biomass.
@wang_biodiversity_2018 have shown that total nutrient uptake might increase
with the length of trophic chains and so results in higher community biomass. While
the predictions of @wang_biodiversity_2018 were restricted to food chains, and
thus might not hold in more complex food-webs with omnivory, our
results are in agreement with their model predictions and highlight the
importance of food chain length as a determinant of community biomass in food-webs.
In addition, mean trophic level had a positive effect on biomass stability by
its negative effect on $\overline{CV_{sp}}$, thereby increasing average
population 
stability. This effect might result from a higher stability of population of higher
trophic level as predicted by food chain models in the presence of strong
top-down control [@shanafelt_stability_2018; @barbier_pyramids_2019]. An
alternative explanation is that higher mean trophic level complexify the biomass
fluxes through the food-webs, which might reduce biomass fluctuations in lower
trophic levels [@mccann_diversitystability_2000].

In contrast, connectance had no effect on community biomass. One might have expected
that an increase in connectance may translate in a higher total nutrient uptake
and so results in higher community biomass in the fish compartment. 
However, some studies suggest that a greater connectance should decrease
complementarity in resource use between species, which is expected to
lower community biomass [@thebault_food-web_2003; @poisot_trophic_2013].

>>>>

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
  word_document:
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
csl: ecology-letters.csl
---

@@

- Number of references: 92
- Number of figures: 3
- Number of tables: 1
- Number of text boxes: 0

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural context. Here we quantify the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardized monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first reveal that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we show that mean trophic level decrease population variability thereby
increasing community stability, and contrary to expectation, a destabilizing
effect of species richness. Finally, we find that environmental conditions
affect both biomass and its stability via effects on richness and network
structure. Strikingly, the variation of stability are mediated by population
stability rather than synchrony, which contrasts with results from
single-trophic communities and might be characteristic of multi-trophic
communities.

@@

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a major effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

@@

Species richness has been found to increase primary productivity, community
biomass and its temporal stability in a diversity of ecosystems and taxa both
empirically and experimentally [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020].  However, these studies mainly focused on
communities composed of a single trophic level, generally primary producers,
thereby ignoring potential effects of food-web structure [but see
@scherber_bottom-up_2010]. Conversely, theoretical works suggest that both
species richness and food-web structure affect ecosystem properties such as
productivity, total biomass and its stability. For example, the presence of
species at higher trophic levels may increase total primary production, through
the selection of primary producers which are larger and more complementary
[@wang_biodiversity_2018]. In communities including producers and consumers, the
connectance which measures the degree of consumer generalism is predicted to
modulate the diversity-productivity relationship [@thebault_food-web_2003;
@thebault_cascading_2007]. Similarly, theoretical studies generally show strong
links between food-web structure and stability [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].  Overall, the evidences from theoretical
model suggests that both species richness and food-web structure should affect
community biomass and stability, but empirical evidences of the effects of
network structure on community biomass and stability are lacking.

While community stability can be measured in many ways in theoretical studies,
most empirical studies measure community stability as the inverse of the
coefficient of variation of community productivity, biomass or abundance, also
called temporal stability [@kefi_advancing_2019]. This stability metric is also
well grounded in theory [e.g. @thebault_trophic_2005; @loreau_species_2008;
@thibaut_understanding_2013] and can be decomposed into two components:
population stability and synchrony [@thibaut_understanding_2013].  Studies
considering a single trophic level [@yachi_biodiversity_1999;
@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020] as
well as theoretical prediction for multitrophic communities
[@thebault_trophic_2005] tend to highlight that asynchrony in population
dynamics as key for the stabilizing effects of species richness on community
biomass. Theory also predicts that higher food-web connectance can
counterbalance the positive effect of species richness on the temporal stability
of community biomass, because it may increase interaction strength and
thereby increases population variability, leading to a negative relationship
between species richness and the temporal stability of populations
[@thebault_trophic_2005]. Predators can also have cascading effects on both
population variability and synchrony at lower trophic levels
[@teng_dynamics_2004; @shanafelt_stability_2018]. Overall, these theoretical
studies show that the effects of species richness and food-web structure, such
as connectance and average trophic level, are potentially conflicting or
synergistic in such a way that they need to be addressed jointly.

The joint study of the effects of species richness and food-web structure on total
biomass and its stability is even more important in the context of environmental
changes. Recent studies comparing natural communities have shown that
environmental characteristics determine in part community biomass
[@grace_integrative_2016; @woods_testing_2020] and its temporal stability
[@franssen_annual_2011; @hansen_climate_2013; @bluthgen_land_2016;
@de_boeck_patterns_2017; @friedpetersen_drivers_2020]. The environment can
affect ecosystem function or stability directly. For example, larger freshwater ecosystems
tend to have more available resources and thus community biomass
[@post_ecosystem_2000; @doi_resource_2009] while agricultural intensification
destabilizes bird communities by decreasing population stability
[@olivier_urbanization_2020]. The effects of environment can also go through
effects on species richness and food-web structure, as larger ecosystems host
food-webs with higher diversity and trophic levels [@mchugh_dual_2010] whereas
temperature can affect food-web diversity and structure in contrasted ways
depending on the ecosystem studied [@ogorman_unexpected_2017;
@gauzens_biodiversity_2020]. Environmental characteristics have thus to be
accounted for if we are to tease apart the effects of species richness and
foodweb structure on community biomass and its temporal stability.

@@

Stream fish communities constitute a good model to study the effects of network
structure on total biomass and its temporal stability because their trophic
interactions are well known,. A substantial body of literature supports that
trophic interactions can be inferred through body size and diet informations
[@jennings_weak_2001; @beckerman_foraging_2006; @gravel_inferring_2013;
@brose_predator_2019; @portalier_mechanics_2019]. Using long-term monitoring of
stream fish communities [@edeline_ecological_2013], we propose one of the first
attempts to test empirically the theoretical expectations on the
diversity-stability relationships according for community, trophic complexity
and environmental characteristics.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between fishes and the resources (Fig.
SXX), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on
plankton while adults may feed on algae and fishes, with bigger individuals
eating bigger fishes. Each fish species was divided in nine body size
classes, each corresponding to a trophic species (appendix B, Fig. S2). A
trophic species constitutes a group of individuals from the same species and
size class, which are expected to share the same sets of prey and predators.
Additionally to the trophic species, we added seven resource nodes present in
the ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Fig. \@ref(fig:matmet)C
& D). The trophic species and the resources constituted the 412 nodes of the
metaweb, i.e. 45 fish species divided into nine body size classes plus seven
resource nodes. The inference of trophic interactions was shown to be more
accurate when considering body size classes, i.e. trophic species, than species
[@jennings_weak_2001].

@@

Species richness and mean trophic level had a significant positive direct effect
on community biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-bm)A, B, C, Table
\@ref(tab:indir-sem)), while connectance had no direct effect.

The two PCA axes synthesising environmental variables had no significant direct
effects on community biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, related to the stream size, was positively linked to community biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Fig. \@ref(fig:sem-bm) F, G, Table \@ref(tab:indir-sem)). The second axis, related to the average
annual temperature was positively related
to community biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Fig. \@ref(fig:sem-bm) G, Table \@ref(tab:indir-sem)). Streams
with higher temperature and BOD contained more species, indirectly increasing
community biomass. The total effect of PCA1, PCA2 and species richness on
community biomass were of the same magnitude.

@@

By its effects on synchrony and population variability ($\overline{CV_{sp}}$),
species richness had a negative total effect on biomass stability ($r_\delta =
`r rich_on_stab$total`$, Table \@ref(tab:indir-sem)). This effect resulted from
two opposite effects. On the one hand, species richness had a direct negative
effect on synchrony ($r_\delta = `r rich_on_sync$direct`$, Fig.
\@ref(fig:sem-stab) A, B), thereby increasing stability ($r_\delta = 
`r rich_on_stab$via_sync`$). On the other hand, species richness had a positive
direct effect on $\overline{CV_{sp}}$ ($r_\delta = `r rich_on_cv_sp$direct`$,
Fig. \@ref(fig:sem-stab) A, C), thereby decreasing biomass stability ($r_\delta
= `r rich_on_stab$via_cv_sp`$, Table \@ref(tab:indir-sem)), which outweighted
the decreasing effect of species richness on synchrony. The negative effect of
$\overline{CV_{sp}}$ on temporal stability was twice higher than the negative
effect of synchrony (resp.  $r_\delta = `r cv_sp_on_stab`$ and $r_\delta = 
`r sync_on_stab`$, Fig.  \@ref(fig:sem-stab)A). Population variability and
synchrony constitute a mathematical decompositon of the stability, contributing
respectively to 1 and 0.5 respectively (Eq. \@ref{eq:stab}), meaning that their
$r_\delta$ difference to temporal stability is due to their standard deviation,
and so that the standard deviation of $\overline{CV_{sp}}$ is of the same
magnitude than the one of synchrony.

The total effect of mean trophic level and species richness on temporal
stability of biomass were of the same magnitude (resp.  
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).  The mean trophic level had only a direct negative effect
on $\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$, Fig.
\@ref(fig:sem-stab)A, F) and thereby had a total positive effect on temporal
stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$, Table
\@ref(tab:indir-sem)). Food webs in which community biomass is located on
average at higher trophic levels thus tend to have higher population and
community stability. Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-stab)A, Table \@ref(tab:indir-sem)).

The total effect of PCA2, related to average annual temperature and BOD, on
temporal stability was of higher magnitude than the effects of species richness
and mean trophic level (Table \@ref(tab:indir-sem)).  The PCA2 had a total
negative effect on temporal stability, meaning that higher temperature and BOD
results in lower stability ($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)).  PCA2 had an indirect positive effect on temporal
stability through a direct positive effect on mean trophic level ($r_\delta = 
`r env_on_stab$via_tlvl["log_RC2"]`$, Fig. \@ref(fig:sem-stab) I, Table
\@ref(tab:indir-sem)). However, PCA2 had a negative effect on temporal stability
through a direct positive effect on $\overline{CV_{sp}}$ and species richness
(resp.  $r_\delta = `r env_on_cv_sp$direct["log_RC2"]`$ and $r_\delta = 
`r env_on_stab$via_richness["log_RC2"]`$, Fig. \@ref(fig:sem-stab)G, Fig.
\@ref(fig:sem-bm)G, Table \@ref(tab:indir-sem)). This was
also true for the PCA1, related to stream size, with a resulting very slight
positive effect of PCA1 on temporal stability ($r_\delta = 
`r env_on_stab$total["log_RC1"]`$, Table \@ref(tab:indir-sem)).  Warmer and
larger streams thus tend to have more species and more variability in population
biomass, then have a lower stability of community biomass.

@@

We found that species richness increased community biomass in line with the
results on plants and tree communities [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018] as well
as those found for fish communities [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020] and for theoretical food-web models
[@schneider_animal_2016]. Theory suggests that the positive effect of species
richness on community biomass might be held by resource use complementarity
[@carey_determining_2011], which is driven by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or by positive selection
effect, which is driven by the dominance of the most productive species
[@loreau_biodiversity_2001]. Recent findings suggest that dominance, a positive
selection effect [@loreau_partitioning_2001], is the main driver of the observed
positive relationship between biomass and species richnes in fish communities
[@maureaud_biodiversityecosystem_2019; @woods_testing_2020].

We found that species richness decreases synchrony, threby increasing stability.
This is in accordance with most studies on experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011] and natural communities
of birds, grasslands, bats and butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020]. It is thought that it is the main effect by which
species richness increases stability [but see @declerck_species_2006;
@valencia_synchrony_2020].  Our result thus extend that species richness
decreases synchrony also in multi-trophic communities.  However, theory does not
clearly state what would be the effect of food-web structure on the relationship
between species richness and synchrony [@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005; @mccann_more_2009].

The positive effect of species richness on population variability, thereby
decreasing stability, is more intriguing. A previous meta-analysis has indeed
reported that species richness had often a negative effect on population
variability in multi-trophic communities while a positive effect is often found
in single-trophic communities [@jiang_different_2009]. In food-webs, theoretical
models predict that the relationship between the temporal stability of populations
and species richness depends on interaction strength and its distribution
[@mccann_reevaluating_1997; @thebault_trophic_2005; @downing_temporal_2020],
with stronger interactions promoting a negative relationship between species
richness and stability. Therefore, it is possible that our results are due to
strong trophic interactions, leading to more fluctuations in population biomass,
but this hypothese was not testable with the available dataset. Alternatively,
other authors suggest that this effect might be due to an increase of
demographic stochasticity with an increase of species richness [@loreau_species_2008].

A major finding of the study is that we found a negative relationship between
species richess and temporal stability, contrast with most experimental and
empirical studies to date [@franssen_annual_2011; @tilman_biodiversity_2006;
@valencia_synchrony_2020; @pennekamp_biodiversity_2018;
@olivier_urbanization_2020], but coincide with a few empirical ones
[@declerck_species_2006]. This finding means that the destabilising effect of
species richness through population variability was superior its stabilising
effect through synchrony. It can happen either because of the relative strength
of the effect species richness on both population variability and synchrony or
because of the overall variability of synchrony and population variability in
the our study system.

We found indeed that the population variability had a standardized effect
on the stability of community biomass twice stronger than synchrony, meaning that the
variability of synchrony was equivalent the one of population variability, in
contrast with a previous study which found that the variability of synchrony was
twice higher than the variability of population variability
[@olivier_urbanization_2020]. Therefore, the low variability of synchrony in our
system might be a cause for the observed negative effect of species richness on
temporal stability of community biomass despite the strong negative effect of species
richness on synchrony. Could the lower variability of synchrony be a
characteristic of multitrophic communities? Theoretical models suggest that
generalist predator can either synchronise or desynchronize the dynamics its
prey, depending on the distribution of interaction strength and competition
between preys [@raimondo_interspecific_2004; @korpimaki_predatorinduced_2005;
@mccann_more_2009]. Alternatively, it is possible that, in aquatic food-webs,
the occurrence of large oscillatory dynamics typical of predator-prey dynamics
leads to a larger variability of population variability than the one of synchrony
[@mccann_reevaluating_1997]. Another hypothese is that food-webs are
characterized by highly uneveness biomass distribution, which can decrease the
variability of synchrony, i.e. decrease the strength of the porfolio effect
[@doak_statistical_1998; @thibaut_understanding_2013].

All together, the destabilising effect of species richness on temporal stability
raises more questions than it solves and then highlight the need for more
investigations about the difference between single and multitrophic
communities in natural communities.

@@

Our analysis further indicates that network structure, and in particular average
trophic level was strongly related to community biomass, as higher total
biomass was found in communities with an higher average trophic level. These results are
in accordance with the predictions of theoretical models [@schneider_animal_2016;
@wang_biodiversity_2018]. @wang_biodiversity_2018 proposed that this effect is driven
by predators selecting primary producers with higher resource use efficiency and predators
enhancing complementarity between primary producers, a mechanism which
drive the positive effect of horizontal diversity on productivity
[@loreau_biodiversity_2001].

In addition, our results show that communities with an higher average trophic
level had a lower population variability, and thereby were more stable. This
result is in accordance with theoretical models [@shanafelt_stability_2018;
@barbier_pyramids_2019; @mccann_diversitystability_2000] and might result from a
higher stability of populations of higher trophic levels
[@shanafelt_stability_2018; @barbier_pyramids_2019] or a damping effect of
higher trophic levels on the variability of lower trophic level populations
[@mccann_diversitystability_2000].

In contrast, we found that connectance was not related to community biomass,
which contradicts theoretical models which suggest that a greater connectance should
decrease complementarity in resource use between species, which is expected to
lower community biomass [@thebault_food-web_2003; @poisot_trophic_2013].
However, the effect of connectance on productivity has been poorly investigated.
